---
title: "Credit Card Point Calculator"
output: html_document
date: "2024-07-31"
author: Lennox Garay
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Chunk 1 ###
This simply loads the necessary packages to run the following code.
If you do not have the following R packages, run Chunk 2 first. 
Run chunks by clicking the green arrow on the top right corner of the {r} line (the first line in each chunk). 

```{r}
library(shiny)
library(shinythemes)
library(ggplot2)

```

### Chunk 2 ### 
DO NOT run this unless Chunk 1 produced errors. 

```{r}
### Please make sure you do not already have these installed. 
### Try to run chunk 1 before running this chunk. 
install.packages("shiny")
install.packages("shinythemes") 
install.packages("ggplot2")
```

Once packages are installed and loaded, you are ready to run the calculator! 

### Chunk 3 ### 

Do not change anything in this chunk. It will certainly break something. Simply run it as before by clicking the green arrow on the first line of the chunk. You can follow the same line as where {r} is.  

```{r}
ui <- fluidPage(
    theme = shinytheme("cerulean"),  # Apply the cerulean theme
  tags$head(
    tags$style(HTML("
      #dataTableContainer {
        position: absolute;
        top: 150px;
        right: 150px;
        width: calc(100% - 150x); /* Adjust according to the sidebar width */
        z-index: 50; /* z-index controls how centered the TABLE output is. */
      }
      #sidebar {
        width: 300px;
        float: right;
        padding-right: 30px;
      }
    "))
  ),
  tags$head(
    tags$style(HTML("
      #dataPlotContainer {
        position: absolute;
        top: 300px;
        right: 5px;
        width: calc(100% - 500px); /* Adjust according to the sidebar width */
        z-index: 20; /* z-index controls how centered the TABLE output is. */
      }"
      ))
  ),
  
  ## App Title 
  titlePanel("Credit Card Point Value Calculator Conditioned on Spend Behavior"),
  
  ### Sidebar layout input definitions 
  sidebarLayout(
    sidebarPanel(
      
      HTML("<h3>Input parameters</h3>"),
      ## Grocery $ 
      numericInput("groceries", label = "Groceries:", min=0, value = 500), ## flights 
      numericInput('restaurants', label = "Restaurants", min=0, value = 300), # food
      
      numericInput("flights", label = "Flights:", min=0, value = 100),
      ## hotel/misc travel 
      numericInput("hotel", label = "Hotel and Misc Travel:", min=0, value = 100), 
      ## other 
      numericInput("other",label =  "Other:", min=0, value = 700),
      
      ## ("update" = actual button command, "Update" = button label)
      actionButton("update", label =  "Update"),
      
      mainPanel(tags$label(h3('Status/Output')), # Status/Output Text Box
                                      verbatimTextOutput('contents'),
                                      tableOutput('tabledata') # Results table
                                    )), # mainPanel()
      helpText("Note: Each category below is the expenditure per month.")
    ),
       # Container for the data frame output
  div(id = "dataTableContainer",
      tableOutput("dataTable")
  ), 
  ## Container for the dataPlot 
  div(id= "dataPlotContainer", 
      plotOutput("dataPlot")
)
) 

server <- function(input, output){
  
  Table <- reactive({
    
    total = input$groceries + input$restaurants + input$flights + input$hotel + input$other 
    amex_pts = 4*input$groceries + 4*input$restaurants + 3*input$flights + input$hotel + input$other
    amex_cash_value = 1*amex_pts/100
    chase_pts = input$groceries + 3*input$restaurants + 3*input$flights + 3*input$hotel + input$other
    chase_cash_value = 1.25*chase_pts/100
    
    
    ## isolate returns the data.frame object while keeping Table() as a function that we will cal in the server function. 
    ## doing it this way will allow us to call the Table() in the UI function so it can be moved absolutely. 
    isolate({
      data.frame("Total Spent" = total, "Amex Points" = amex_pts,
                     "Amex Cash Value" = amex_cash_value, "Chase Points" = chase_pts,
                     "Chase Cash Value" = chase_cash_value)
    })
  })


CashTable = reactive({
  chase_points_to_cash = function(){
  chase_pts = input$groceries + 3*input$restaurants + 3*input$flights + 3*input$hotel + input$other
  
  
  x = rep(1:chase_pts)
  cash_value = numeric(chase_pts)
  for(i in 1:chase_pts){
  cash_value[i] = 1.25*x[i]/100
  }
  list(chase_y_plot = cash_value)
}

  amex_points_to_cash = function(){
  amex_pts = 4*input$groceries + 4*input$restaurants + 3*input$flights + input$hotel + input$other
  
  x = rep(1:amex_pts)
  cash_value2 = numeric(amex_pts)
  for(i in 1:amex_pts){
  cash_value2[i] = 1*x[i]/100
  }
  list(amex_y_plot = cash_value2)
}

## chase_cash_value as object from the function 
temp = chase_points_to_cash()
chase_value = temp$chase_y_plot 


temp2 = amex_points_to_cash()
amex_value = temp2$amex_y_plot

df1 = data.frame("Chase_Value" = chase_value)
df2 = data.frame("Amex_Value" = amex_value)

## additional blocks for plot extensions
e1 <- e2 <- rep(0, length(df1$Chase_Value))

# Vectorized conditional logic

if(df1$Chase_Value[length(df1$Chase_Value)] < df2$Amex_Value[length(df2$Amex_Value)]){
  chase_max = df1$Chase_Value[length(df1$Chase_Value)]
  x = seq(1:length(df2$Amex_Value))
  diff = length(df2$Amex_Value) - length(df1$Chase_Value)
  y1 = df1$Chase_Value
  y2 = rep(chase_max,diff)
  y = c(y1,y2)
  df1 = data.frame("Chase_Value" = y)
} else{  
    amex_max = df2$Amex_Value[length(df2$Amex_Value)]
    x = seq(1:length(df1$Chase_Value))
    diff = length(df1$Chase_Value) - length(df2$Amex_Value)
    y1 = df2$Amex_Value
    y2 = rep(amex_max,diff)
    y = c(y1,y2)
    df2 = data.frame("Amex_Value" = y)
  }



isolate({
  ggplot() + geom_line(data = df1, mapping = aes(x = seq(from = 0, to =length(Chase_Value)-1)
                                               , y=Chase_Value, color = 'Chase')) + 
  geom_line(data = df2, mapping = aes(x = seq(from = 0, to=length(Amex_Value)-1),
                                      y = Amex_Value, color = 'Amex')) +
  labs(title = "Chase vs Amex Points Cash Value Conditioned on Spend Behavior", color = "Credit Card Company") +
  xlab("Points") + ylab("Cash Value")
})

})
## NEED TO MAKE THE PLOTSSS
  
  ## Render data frame in the UI so that I can call it using tags$head #Table. 
  
  output$dataTable = renderTable({
    Table()
  })
  
  output$dataPlot <- renderPlot({
    CashTable()
  })
  
  
} 

shinyApp(ui = ui, server = server)

```
